---
title: "R Markdown of the Manuscript: Environmental drivers of the resistome across the Baltic Sea"
author: "Joeselle Serrana"
date: "2024-03-27"
output: html_document
editor_options: 
  chunk_output_type: console
---

### Step 1A: Data Preparation - Load required library and set working directory 

```{r}
# Load libraries for data manipulation, analysis, and visualization
library("phyloseq")      # Tools for microbiome census data
library("microeco")      # Ecological analysis of microbiome data; visit https://github.com/ChiLiubio/microeco for more details
packageVersion("microeco")

# Load libraries for network analysis
library("meconetcomp")   # Tools for microbial ecological network comparison
library("ggraph")        # Grammar of graphics for relational data
library("caret")         # Classification and regression training

# Load libraries for general and specialized visualizations
library("ggplot2")       # Data visualization
# library("ggradar")     # Radar charts
# library("ggcor")       # Correlation analysis
library("microViz")      # Microbiome data visualization
library("car")           # Companion to Applied Regression
library("magrittr")      # Pipe-friendly tools for better syntax
library("ComplexHeatmap")# Advanced heatmap visualization
library("aplot")         # Annotate or arrange multiple plots

# Load additional libraries for specialized plotting and data manipulation
library("ggh4x")         # Extensions for ggplot2
library("RColorBrewer")  # Color palettes
library("dplyr")         # Data manipulation
library("mice")          # Multivariate Imputation by Chained Equations
library("textshape")     # Text shaping tools
library("readxl")        # Read Excel files

setwd("~/Library/Mobile Documents/com~apple~CloudDocs/Documents/(In Revision) Baltic Sea AMR/MS-Baltic-Sea-AMR_Analysis")
#setwd("C:/Users/JoeselleSerrana/iCloudDrive/Documents/(In Revision) Baltic Sea AMR/MS-Baltic-Sea-AMR_Analysis")
```

### Step 1B: Data Preparation - Generate phyloseq and microtable objects from input data 

```{r}
### STEP 1a: Generate phyloseq object for microtable community dataset
### A phyloseq object contains OTU table (taxa abundances), sample metadata, taxonomy table (mapping between OTUs and higher-level taxonomic classifications), and phylogenetic tree (relations between the taxa).

# Import sample data from Excel file
metadata <- read_excel("data/data.xlsx", sheet = "metadata")
metadata <- column_to_rownames(metadata, loc = 1)
metadata <- sample_data(as.data.frame(metadata[, 2:18]))

# Transform environmental variables to meet normality and homogeneity of variance conditions
metadata.log <- metadata
metadata.log[, 7:15] <- log1p(abs(metadata.log[, 7:15]))

# Calculate the variation inflation factors (VIF) of the environmental variables using Salinity as the response variable
metadata.vif <- lm(Salinity ~ ., data = abs(metadata.log[, 7:15]))
summary(metadata.vif)
vif.values <- vif(metadata.vif)
vif.values

# Remove variables with VIF greater than 10 from metadata (if needed)
# vif.index <- t(as.data.frame(vif.values))
# vif.index <- as.data.frame(subset(vif.values, vif.values > 5))
# vif.index <- rownames(vif.index)
# `%ni%` <- Negate(`%in%`)
# metadata.fin <- subset(metadata.log, select = names(metadata.log) %ni% vif.index)
# vif.index

# Import counts annotation data from Excel file and define row names
otu.tbl <- read_excel("data/data.xlsx", sheet = "motus")
otu.tbl <- column_to_rownames(otu.tbl, loc = 1)
arg.tbl <- read_excel("data/data.xlsx", sheet = "arg")
arg.tbl <- column_to_rownames(arg.tbl, loc = 1)
mrg.tbl <- read_excel("data/data.xlsx", sheet = "mrg")
mrg.tbl <- column_to_rownames(mrg.tbl, loc = 1)
mge.tbl <- read_excel("data/data.xlsx", sheet = "mge")
mge.tbl <- column_to_rownames(mge.tbl, loc = 1)
vfg.tbl <- read_excel("data/data.xlsx", sheet = "vfg")
vfg.tbl <- column_to_rownames(vfg.tbl, loc = 1)

# Remove decimals from columns in data frames
arg.tbl[1:59] <- lapply(arg.tbl, as.integer)
mrg.tbl[1:59] <- lapply(mrg.tbl, as.integer)
mge.tbl[1:59] <- lapply(mge.tbl, as.integer)
vfg.tbl[1:59] <- lapply(vfg.tbl, as.integer)

### STEP 1b: Create phyloseq object
otu.psq <- phyloseq(
  otu_table(as.matrix(otu.tbl[1:59]), taxa_are_rows = TRUE),
  tax_table(as.matrix(otu.tbl[60:66])),
  sample_data(metadata.log)
)
arg.psq <- phyloseq(
  otu_table(as.matrix(arg.tbl[1:59]), taxa_are_rows = TRUE),
  tax_table(as.matrix(arg.tbl[60:63])),
  sample_data(metadata.log)
)
mrg.psq <- phyloseq(
  otu_table(as.matrix(mrg.tbl[1:59]), taxa_are_rows = TRUE),
  tax_table(as.matrix(mrg.tbl[60:63])),
  sample_data(metadata.log)
)
mge.psq <- phyloseq(
  otu_table(as.matrix(mge.tbl[1:59]), taxa_are_rows = TRUE),
  tax_table(as.matrix(mge.tbl[60:65])),
  sample_data(metadata.log)
)
vfg.psq <- phyloseq(
  otu_table(as.matrix(vfg.tbl[1:59]), taxa_are_rows = TRUE),
  tax_table(as.matrix(vfg.tbl[60:63])),
  sample_data(metadata.log)
)

### STEP 1c: Create microtable object for microeco downstream analyses

# Create microeco object
otu.mco <- file2meco::phyloseq2meco(otu.psq, auto_tidy = FALSE)
otu.mco$tidy_dataset()
arg.mco <- file2meco::phyloseq2meco(arg.psq, auto_tidy = FALSE)
arg.mco$tidy_dataset()
mrg.mco <- file2meco::phyloseq2meco(mrg.psq, auto_tidy = FALSE)
mrg.mco$tidy_dataset()
mge.mco <- file2meco::phyloseq2meco(mge.psq, auto_tidy = FALSE)
mge.mco$tidy_dataset()
vfg.mco <- file2meco::phyloseq2meco(vfg.psq, auto_tidy = FALSE)
vfg.mco$tidy_dataset()

### STEP 1d: Add abundance and beta diversity estimates

# Calculate the taxa abundance at each taxonomic rank using cal_abund()
otu.mco$cal_abund()
arg.mco$cal_abund()
mrg.mco$cal_abund()
mge.mco$cal_abund()
vfg.mco$cal_abund()

# Calculate alpha diversity
otu.mco$cal_alphadiv(PD = FALSE)
arg.mco$cal_alphadiv(PD = FALSE)
mrg.mco$cal_alphadiv(PD = FALSE)
mge.mco$cal_alphadiv(PD = FALSE)
vfg.mco$cal_alphadiv(PD = FALSE)

# Calculate beta diversity
otu.mco$cal_betadiv(unifrac = FALSE)
arg.mco$cal_betadiv(unifrac = FALSE)
mrg.mco$cal_betadiv(unifrac = FALSE)
mge.mco$cal_betadiv(unifrac = FALSE)
vfg.mco$cal_betadiv(unifrac = FALSE)

# Save files as RDS
saveRDS(otu.psq, "out/otu.psq.RDS")
saveRDS(arg.psq, "out/arg.psq.RDS")
saveRDS(mrg.psq, "out/mrg.psq.RDS")
saveRDS(mge.psq, "out/mge.psq.RDS")
saveRDS(vfg.psq, "out/vfg.psq.RDS")
saveRDS(otu.mco, "out/otu.mco.RDS")
saveRDS(arg.mco, "out/arg.mco.RDS")
saveRDS(mrg.mco, "out/mrg.mco.RDS")
saveRDS(mge.mco, "out/mge.mco.RDS")
saveRDS(vfg.mco, "out/vfg.mco.RDS")
```

### Step 2A: Visualization - Generate site map 

```{r}
# Load necessary libraries
library(ggOceanMaps)
library(ggspatial)
library(RColorBrewer)

# Prepare metadata2
metadata2 <- metadata[order(metadata$Sample, decreasing = FALSE), ]
metadata2$Sample <- gsub("S", "", as.character(metadata2$Sample))
metadata2$Region <- factor(metadata2$Region, levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", 
                                                        "Dead Zone (Sörmland)", "Östergötland", 
                                                        "Dead Zone (Mid-South)", "Southern Baltic"))

# Create the basemap and plot
basemap(limits = c(12, 20, 53, 67), 
        bathymetry = TRUE, glaciers = FALSE, land.col = "grey75", 
        land.border.col = "grey50",
        bathy.style = "raster_continuous_blues",
        legends = TRUE) +
  ggspatial::geom_spatial_point(data = metadata2, aes(x = Longitude, y = Latitude, color = Region), size = 3, alpha = 0.9) +
  ggspatial::geom_spatial_text_repel(data = metadata2, aes(x = Longitude, y = Latitude, label = Sample, color = Region), max.overlaps = Inf, size = 3) +
  scale_color_brewer(palette = "Dark2") +
  ggspatial::annotation_scale(location = "br", width_hint = 0.5) +
  ggspatial::annotation_north_arrow(location = "tr", which_north = "true", style = north_arrow_fancy_orienteering) +
  labs(x = NULL, y = NULL) +  # Remove axis titles
  theme_bw() +  # Apply a clean theme
  theme(
    axis.text = element_text(size = 8),  # Resize axis labels
    axis.title = element_blank(),  # Remove axis titles
    axis.text.y = element_text(size = 8, angle = 90)  # Rotate y-axis label
  )
```

```{r, graphical abstract map}
# Load necessary libraries
library(ggOceanMaps)
library(ggspatial)
library(RColorBrewer)

# Create the basemap and plot
basemap(limits = c(5, 50, 52, 67), bathymetry = TRUE, glaciers = FALSE, land.col = "grey75", land.border.col = "grey75", bathy.style = "raster_continuous_blues",legends = TRUE)
```

```{r}
library(ggOceanMaps)
library(sf)
library(ggplot2)

# Define the latitude and longitude ranges for Europe
lon_range <- c(-30, 60)
lat_range <- c(30, 80)

# Create the base map
europe_map <- basemap(limits = c(lon_range, lat_range), land.col = "grey75", land.border.col = "grey50")

# Display the customized map
print(europe_map)
```

### Step 2B: Visualization - Environmnetal variables and autocorrelation 

```{r}
# Clone the dataset and redefine Region levels
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", 
                                                      "Dead Zone (Sörmland)", "Östergötland", 
                                                      "Dead Zone (Mid-South)", "Southern Baltic"))

# Create a trans_env object and add environmental data
t1 <- trans_env$new(dataset = dataset, env_cols = 7:15, complete_na = TRUE)

# Perform ANOVA to test significance of variables across regions
t1$cal_diff(group = "Region", method = "anova")
head(t1$res_diff)

# Create boxplot plots for each environmental variable
tmp <- list()
for (i in colnames(t1$data_env)) {
  tmp[[i]] <- t1$plot_diff(measure = i, add_sig_text_size = 3, xtext_size = 8) +
    geom_boxplot(outlier.shape = NA, fill = RColorBrewer::brewer.pal(8, "Dark2"), color = NA, alpha = 0.5) +
    theme_bw() +
    theme(
      axis.title.y = element_text(face = "bold"),
      axis.text.y = element_text(angle = 90, vjust = 0.5, hjust = 0.5, size = 8),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank(),
      legend.position = "none",
      panel.grid = element_blank(),
      plot.margin = unit(c(0.1, 0, 0, 1), "cm")
    )
}

# Arrange and display boxplot plots in a grid
plot(gridExtra::arrangeGrob(grobs = tmp, ncol = 3))

# Calculate autocorrelations among variables
t1$cal_autocor()  # By default, calculates autocorrelation among all variables

# Optionally specify a grouping variable (e.g., Region) for autocorrelation analysis
# t1$cal_autocor(group = "Region")
# This line would calculate autocorrelation for each region separately, if needed
```

### Step 2C: Visualization - Relative abundance and frequency composition 

```{r, Visualize composition}
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", ntaxa = 8, groupmean = "Region")
t1$plot_donut(label = FALSE, color_values = c(brewer.pal(n = 8, name = "Dark2"), "gray90")) + theme(legend.position = "right")
t1$plot_radar(values.radar = c("0%", "25%", "50%"), grid.min = 0, grid.mid = 0.25, grid.max = 0.5, color_values = c(brewer.pal(n = 5, name = "RdYlBu"), "gray90"))
 
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_norm$new(dataset = dataset)
dataset.pa <- t1$norm(method = "pa"); dataset.pa

t1 <- trans_abund$new(dataset = dataset.pa, taxrank = "Type", ntaxa = 8, groupmean = "Region")
F0 <- t1$plot_line(position = position_dodge(0.3), xtext_angle = 0, color_values = c(rev(brewer.pal(n = 8, name = "Spectral")), "gray90"), errorbar_width = 1, errorbar_size = 0.25,
                   point_alpha = 1, line_alpha = 0.5, point_size = 1.5) + theme_bw() + 
      theme(axis.text.x = element_blank(), panel.grid = element_blank(), legend.position = "none", axis.ticks.x = element_blank()) +
      xlab("North to South across 1,150 km distance") + ylab("Frequency of \n Top 8 Predicted ARGs (%)") #+ ggtitle("Top 8 ARGs")
F0
 
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = rev(c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic")))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", ntaxa = 8, groupmean = "Region")
C0 <- t1$plot_bar(coord_flip = FALSE, color_values = brewer.pal(n = 8, name = "RdYlBu")) +
      theme_bw() + theme(legend.position = "right", axis.title.x = element_text(size = 10), axis.ticks.y = element_blank(),  #axis.text.x = element_text(angle = 25, hjust = 1), 
                         axis.line.y = element_blank()) +
      ylab("ARG Relative abundance (%)") 
C0

C0 <- t1$plot_bar(clustering_plot = TRUE, color_values = rev(brewer.pal(n = 8, name = "Spectral")))
C0[[1]] <- C0[[1]] + theme_bw() + theme(legend.position = "none", axis.title.x = element_text(size = 10), axis.ticks.y = element_blank(), axis.line.y = element_blank()) +
      ylab("ARG Relative abundance (%)")
C0
```

```{r, Visualize composition}
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", ntaxa = 8)
# require ggalluvial package
# use_alluvium = TRUE make the alluvial plot, clustering =TRUE can be used to reorder the samples by clustering
# bar_type = FALSE can remove 'others'
C1 <- t1$plot_bar(others_color = "grey90", xtext_keep = FALSE, use_alluvium = TRUE, legend_text_italic = FALSE, xtext_angle = 90, xtext_size = 10, color_values = brewer.pal(n = 8, name = "Spectral")) +
                  ggalluvial::geom_alluvium(width = 1/5, alpha = 0.9) + theme_bw() + xlab("km distance") + ylab("ARG Relative abundance (%)") +
                  theme(legend.position = "right", axis.text.x = element_blank(), axis.ticks.x = element_blank())
C1
 
dataset <- clone(otu.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Phylum", ntaxa = 8)
# require ggalluvial package
# use_alluvium = TRUE make the alluvial plot, clustering =TRUE can be used to reorder the samples by clustering
# bar_type = FALSE can remove 'others'
C2 <- t1$plot_bar(others_color = "grey90", xtext_keep = FALSE, use_alluvium = TRUE, legend_text_italic = FALSE, xtext_angle = 90, xtext_size = 10, color_values = rev(brewer.pal(n = 8, name = "Spectral"))) +
                  ggalluvial::geom_alluvium(width = 1/5, alpha = 0.9) + theme_bw() + xlab("km distance") + ylab("Phylum Relative abundance (%)") +
                  theme(legend.position = "right", axis.text.x = element_blank(), axis.ticks.x = element_blank())
C2

dataset <- clone(mrg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Resistance", ntaxa = 8)
# require ggalluvial package
# use_alluvium = TRUE make the alluvial plot, clustering =TRUE can be used to reorder the samples by clustering
# bar_type = FALSE can remove 'others'
C3 <- t1$plot_bar(others_color = "grey90", xtext_keep = FALSE, use_alluvium = TRUE, legend_text_italic = FALSE, xtext_angle = 90, xtext_size = 10, color_values = brewer.pal(n = 8, name = "YlGnBu")) +
                  ggalluvial::geom_alluvium(width = 1/5, alpha = 0.9) + theme_bw() + xlab("km distance") + ylab("MRG Relative abundance (%)") +
                  theme(legend.position = "right", axis.text.x = element_blank(), axis.ticks.x = element_blank())
C3

dataset <- clone(mge.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Category", ntaxa = 5)
# require ggalluvial package
# use_alluvium = TRUE make the alluvial plot, clustering =TRUE can be used to reorder the samples by clustering
# bar_type = FALSE can remove 'others'
C4 <- t1$plot_bar(others_color = "grey90", xtext_keep = FALSE, use_alluvium = TRUE, legend_text_italic = FALSE, xtext_angle = 90, xtext_size = 10, color_values = rev(brewer.pal(n = 5, name = "YlOrBr"))) +
                  ggalluvial::geom_alluvium(width = 1/5, alpha = 0.9) + theme_bw() + xlab("km distance") + ylab("MGE Relative abundance (%)") +
                  theme(legend.position = "right", axis.text.x = element_blank(), axis.ticks.x = element_blank())
C4

dataset <- clone(vfg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Category", ntaxa = 8)
# require ggalluvial package
# use_alluvium = TRUE make the alluvial plot, clustering =TRUE can be used to reorder the samples by clustering
# bar_type = FALSE can remove 'others'
C5 <- t1$plot_bar(others_color = "grey90", xtext_keep = FALSE, use_alluvium = TRUE, legend_text_italic = FALSE, xtext_angle = 90, xtext_size = 10, color_values = rev(brewer.pal(n = 8, name = "PuBuGn"))) +
                  ggalluvial::geom_alluvium(width = 1/5, alpha = 0.9) + theme_bw() + xlab("km distance") + ylab("VFG Relative abundance (%)") +
                  theme(legend.position = "right", axis.text.x = element_blank(), axis.ticks.x = element_blank())
C5

C1 / C2 
C3 / C4 / C5
```

```{r, Donut plot}
# Load necessary libraries
library(ggplot2)
library(dplyr)
library(scales)

# Define the RPKM values for each category
rpkm_values <- data.frame(
  Type = c("Multidrug", "Glycopeptide", "Beta-lactam", "Bacitracin", "MLS", "Aminoglycoside", "Tetracycline", "Fosmidomycin", "Others"),
  RPKM = c(17767205, 7584706, 3663148, 3540029, 2081478, 1706843, 1641140, 845100, 7550950)
)

# Calculate the total RPKM
total_rpkm <- sum(rpkm_values$RPKM)

# Calculate the percentage contribution of each type
rpkm_values <- rpkm_values %>%
  mutate(Percentage = RPKM / total_rpkm * 100)

# Arrange the data by most abundant to least, but ensure "Others" is last
rpkm_values <- rpkm_values %>%
  arrange(desc(RPKM)) %>%
  mutate(Type = factor(Type, levels = c(as.character(Type[Type != "Others"]), "Others")))

# Define the colors for each category, using reverse Spectral for all except "Others" which is gray90
colors <- c(rev(brewer_pal(palette = "Spectral")(8)), "gray90")

# Create the donut plot
ggplot(rpkm_values, aes(x = 2, y = Percentage, fill = Type)) +
  geom_bar(stat = "identity", width = 1, color = "white") +
  coord_polar(theta = "y") +
  scale_fill_manual(values = colors) +
  xlim(0.5, 2.5) +
  theme_void() +
  theme(legend.position = "right") +
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, Visualize relative abundance of ARG}
# Load necessary libraries
library(ggplot2)
library(agricolae)  # for Duncan.test
library(dplyr)       # for data manipulation
library(ggpubr)

dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Type"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment

# Show 8 taxa at Type level
t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", use_percentage = TRUE)
R1 <- t1$plot_box(group = "Region") + theme_bw() + theme(#axis.text.x = element_blank(),
                                                         axis.text.x = element_text(angle = 90, hjust = 1),
                                                         panel.grid = element_blank(), legend.position = "none"); R1
#t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", use_percentage = TRUE, input_taxaname = "Multidrug")
#R2 <- t1$plot_box(group = "Region") + theme_bw() + theme(axis.text.x = element_blank(), panel.grid = element_blank(), legend.position = "none"); R2
#t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", use_percentage = TRUE, input_taxaname = "Glycopeptide")
#R3 <- t1$plot_box(group = "Region") + theme_bw() + theme(axis.text.x = element_blank(), panel.grid = element_blank(), legend.position = "none"); R3

multi <- as.data.frame(t1$data_abund) %>% filter(grepl("Multidrug", Taxonomy, ignore.case = TRUE))
anova_result_1 <- aov(Abundance ~ Region, data = multi); summary(anova_result_1) # Perform ANOVA
duncan_result_1 <- duncan.test(anova_result_1, "Region", console = TRUE); duncan_result_1 # Perform Duncan's test for multiple comparisons

R2 <- ggboxplot(multi, x = "Region", y = "Abundance", color = "Region", add = "jitter", legend = "none", palette = "Dark2") +  
      geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.85, outlier.shape = NA) + ylim(28, 48) +
      geom_hline(yintercept = mean(multi$Abundance), linetype = 2) + # Add horizontal line at base mean
      #stat_compare_means(method = "anova", label.y = 50) +  # Add global annova p-value
      #stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.", hide.ns = TRUE) + # Pairwise comparison against all
      labs(title = "Multidrug Resistance Genes", y = "RPKM Relative Abundance (%)") + theme_bw() + 
      #theme(axis.text.x = element_text(angle = 35, hjust = 1), axis.title.x = element_blank(), panel.grid = element_blank(), legend.position = "none")
      theme(plot.title = element_text(size = 14), axis.text.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1, size = 12),
            panel.grid = element_blank(), legend.position = "none")

glyco <- as.data.frame(t1$data_abund) %>% filter(grepl("Glycopeptide", Taxonomy, ignore.case = TRUE))
anova_result_2 <- aov(Abundance ~ Region, data = glyco); summary(anova_result_2) # Perform ANOVA
duncan_result_2 <- duncan.test(anova_result_2, "Region", console = TRUE); duncan_result_2 # Perform Duncan's test for multiple comparisons

R3 <- ggboxplot(glyco, x = "Region", y = "Abundance", color = "Region", add = "jitter", legend = "none", palette = "Dark2") +  
      geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.85, outlier.shape = NA) + ylim(12, 26) +
      geom_hline(yintercept = mean(glyco$Abundance), linetype = 2) + # Add horizontal line at base mean
      #stat_compare_means(method = "anova", label.y = 30) +  # Add global annova p-value
      #stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.", hide.ns = TRUE) + # Pairwise comparison against all
      labs(title = "Glycopeptide Resistance Genes", y = "RPKM Relative Abundance (%)") + theme_bw() +
      theme(plot.title = element_text(size = 14), axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.title.x = element_blank(),  axis.text.y = element_text(angle = 90, hjust = 1, size = 12),
            panel.grid = element_blank(), legend.position = "none")

R2 / R3
```

```{r}
## Subset dataset by group.
multi.mco <- clone(arg.mco)
multi.mco$tax_table <- subset(multi.mco$tax_table, Type == "t__Multidrug")
multi.mco$tidy_dataset()
multi.mco

multi.mco$cal_abund(); class(multi.mco$taxa_abund) ## Calculate the taxa abundance at each taxonomic rank using cal_abund().
multi.mco$cal_alphadiv(PD = FALSE); class(multi.mco$alpha_diversity) ## Calculate the Alpha diversity w/ Faith's phylogenetic diversity.
multi.mco$cal_betadiv(unifrac = FALSE); class(multi.mco$beta_diversity) ## Create trans_beta object and measure parameter - dataset$beta_diversity

colors <- c(scales::brewer_pal(palette = "Set1")(8), "gray90")

dataset <- clone(multi.mco)
dataset$sample_table$Region %<>% factor(., levels = rev(c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic")))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Subtype", ntaxa = 8, groupmean = "Region")
C0 <- t1$plot_bar(coord_flip = FALSE, color_values = colors) +
      theme_bw() + theme(legend.position = "right", axis.title.x = element_text(size = 10), axis.ticks.y = element_blank(),  #axis.text.x = element_text(angle = 25, hjust = 1), 
                         axis.line.y = element_blank()) + ylab("Predicted Multidrug ARG Relative abundance (%)") 

C0 <- t1$plot_bar(clustering_plot = TRUE, color_values = colors)
C0[[1]] <- C0[[1]] + theme_bw() + theme(legend.position = "right", axis.title.x = element_text(size = 10), axis.ticks.y = element_blank(), axis.line.y = element_blank()) + ylab("ARG Relative abundance (%)")
C0

dataset$sample_table$ALL <- "ALL"
t1 <- trans_abund$new(dataset = dataset, taxrank = "Subtype", ntaxa = 8, groupmean = "ALL")
P0 <- t1$plot_donut(label = TRUE, color_values = colors)
P0
```

```{r, Visualize relative abundance of MRG}
# Load necessary libraries
library(ggplot2)
library(agricolae)  # for Duncan.test
library(dplyr)       # for data manipulation
library(ggpubr)

dataset <- clone(mrg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Type"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment

# Show 8 taxa at Type level
t1 <- trans_abund$new(dataset = dataset, taxrank = "Resistance", use_percentage = TRUE)
R1 <- t1$plot_box(group = "Region") + theme_bw() + theme(#axis.text.x = element_blank(),
                                                         axis.text.x = element_text(angle = 45, hjust = 1),
                                                         panel.grid = element_blank(), legend.position = "none"); R1
#t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", use_percentage = TRUE, input_taxaname = "Multidrug")
#R2 <- t1$plot_box(group = "Region") + theme_bw() + theme(axis.text.x = element_blank(), panel.grid = element_blank(), legend.position = "none"); R2
#t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", use_percentage = TRUE, input_taxaname = "Glycopeptide")
#R3 <- t1$plot_box(group = "Region") + theme_bw() + theme(axis.text.x = element_blank(), panel.grid = element_blank(), legend.position = "none"); R3

multi <- as.data.frame(t1$data_abund) %>% filter(grepl("arsB", Taxonomy, ignore.case = TRUE))
anova_result_1 <- aov(Abundance ~ Region, data = multi); summary(anova_result_1) # Perform ANOVA
duncan_result_1 <- duncan.test(anova_result_1, "Region", console = TRUE); duncan_result_1 # Perform Duncan's test for multiple comparisons

R2 <- ggboxplot(multi, x = "Region", y = "Abundance", color = "Region", add = "jitter", legend = "none", palette = "Dark2") +  
      geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.85, outlier.shape = NA) + #ylim(28, 48) +
      geom_hline(yintercept = mean(multi$Abundance), linetype = 2) + # Add horizontal line at base mean
      #stat_compare_means(method = "anova", label.y = 50) +  # Add global annova p-value
      #stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.", hide.ns = TRUE) + # Pairwise comparison against all
      labs(title = "arsenic resistance protein ArsB", y = "RPKM Relative Abundance (%)") + theme_bw() + 
      #theme(axis.text.x = element_text(angle = 35, hjust = 1), axis.title.x = element_blank(), panel.grid = element_blank(), legend.position = "none")
      theme(plot.title = element_text(size = 14), axis.text.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1, size = 12),
            panel.grid = element_blank(), legend.position = "none")

multi <- as.data.frame(t1$data_abund) %>% filter(grepl("actP", Taxonomy, ignore.case = TRUE))
anova_result_1 <- aov(Abundance ~ Region, data = multi); summary(anova_result_1) # Perform ANOVA
duncan_result_1 <- duncan.test(anova_result_1, "Region", console = TRUE); duncan_result_1 # Perform Duncan's test for multiple comparisons

R3 <- ggboxplot(multi, x = "Region", y = "Abundance", color = "Region", add = "jitter", legend = "none", palette = "Dark2") +  
      geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.85, outlier.shape = NA) + #ylim(28, 48) +
      geom_hline(yintercept = mean(multi$Abundance), linetype = 2) + # Add horizontal line at base mean
      #stat_compare_means(method = "anova", label.y = 50) +  # Add global annova p-value
      #stat_compare_means(label = "p.signif", method = "t.test", ref.group = ".all.", hide.ns = TRUE) + # Pairwise comparison against all
      labs(title = "copper-translocating P-type ATPase", y = "RPKM Relative Abundance (%)") + theme_bw() + 
      #theme(axis.text.x = element_text(angle = 35, hjust = 1), axis.title.x = element_blank(), panel.grid = element_blank(), legend.position = "none")
      theme(plot.title = element_text(size = 14), axis.text.x = element_blank(), axis.title.x = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1, size = 12),
            panel.grid = element_blank(), legend.position = "none")

R2 + R3
```

### Step 3A: Statistical Analysis - Random forest with differential abundance test  

```{r, type}
# Set seed for reproducibility
set.seed(12345)

# Clone the dataset and redefine Region levels
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", 
                                                      "Dead Zone (Sörmland)", "Östergötland", 
                                                      "Dead Zone (Mid-South)", "Southern Baltic"))

# Create trans_diff object for random forest analysis
# Use 'Subtype' level of taxonomy and no bootstrapping (nresam = 1, boots = 1)
t1 <- trans_diff$new(dataset = dataset, method = "rf", group = "Region", taxa_level = "Type", nresam = 1, boots = 1)

# View the results (MeanDecreaseGini values)
t1$res_diff

# Plot MeanDecreaseGini bar plot
R1 <- t1$plot_diff_bar(use_number = 1:15, keep_prefix = FALSE, keep_full_name = TRUE,
                       group_order = c("Bothnian Bay", "Bothnian Sea", "Dead Zone (Sörmland)", "Östergötland", "Southern Baltic"),
                       color_values = c("#1B9E77", "#D95F02", "#66A61E", "#E6AB02", "#666666")) +
                       theme_bw() +
                       theme(legend.position = "none", panel.grid = element_blank(), axis.title.x = element_text(size = 10)) + ylab("Mean Decrease Gini")

# Display the plot
R1

# Write the results to a TSV file
write.table(t1$res_diff, "out/random_forest.tsv")
```

```{r, subtype}
# Set seed for reproducibility
set.seed(12345)

# Clone the dataset and redefine Region levels
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", 
                                                      "Dead Zone (Sörmland)", "Östergötland", 
                                                      "Dead Zone (Mid-South)", "Southern Baltic"))

# Create trans_diff object for random forest analysis
# Use 'Subtype' level of taxonomy and no bootstrapping (nresam = 1, boots = 1)
t1 <- trans_diff$new(dataset = dataset, method = "rf", group = "Region", taxa_level = "Subtype", nresam = 1, boots = 1)

# View the results (MeanDecreaseGini values)
t1$res_diff

# Plot MeanDecreaseGini bar plot
R2 <- t1$plot_diff_bar(use_number = 1:15, keep_prefix = FALSE, keep_full_name = TRUE,
                       group_order = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)", "Southern Baltic"),
                       color_values = c("#1B9E77", "#D95F02", "#7570B3", "#66A61E", "#E6AB02", "#666666")) +
                       theme_bw() +
                       theme(legend.position = "none", panel.grid = element_blank(), axis.title.x = element_text(size = 10))

# Display the plot
R2

# Write the results to a TSV file
write.table(t1$res_diff, "out/random_forest.tsv")
```

```{r, species}
# Set seed for reproducibility
set.seed(12345)

# Clone the dataset and redefine Region levels
dataset <- clone(otu.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", 
                                                      "Dead Zone (Sörmland)", "Östergötland", 
                                                      "Dead Zone (Mid-South)", "Southern Baltic"))

# Create trans_diff object for random forest analysis
# Use 'Subtype' level of taxonomy and no bootstrapping (nresam = 1, boots = 1)
t1 <- trans_diff$new(dataset = dataset, method = "rf", group = "Region", taxa_level = "Species", nresam = 1, boots = 1)

# View the results (MeanDecreaseGini values)
t1$res_diff

# Plot MeanDecreaseGini bar plot
R2 <- t1$plot_diff_bar(use_number = 1:47, keep_prefix = TRUE, keep_full_name = TRUE,
                       group_order = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)", "Southern Baltic"),
                       color_values = RColorBrewer::brewer.pal(8, "Dark2")) +
                       theme_bw() +
                       theme(legend.position = "none", panel.grid = element_blank(), axis.title.x = element_text(size = 10))

# Display the plot
R2

# Write the results to a TSV file
write.table(t1$res_diff, "out/random_forest.tsv")
```

### Step 3B: Statistical Analysis - Alpha diversity estimation

```{r, levene's test}
# Load required packages
library(phyloseq)
library(car)
library(dplyr)

# Use your phyloseq object
psq <- arg.psq

# Clone the dataset
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

# Function to perform Levene's test for a given diversity metric
perform_levenes_test <- function(diversity_metric, psq, dataset) {
  alpha_diversity_metric <- log(dataset$alpha_diversity[[diversity_metric]])  # Apply log transformation
  diversity_df <- data.frame(alpha_diversity_metric)
  diversity_df$Region <- sample_data(psq)$Region
  
  levene_results <- leveneTest(diversity_df[, 1] ~ Region, data = diversity_df) 
  
  return(levene_results)
}

# Perform Levene's test for Chao1 richness
levene_results_chao1 <- perform_levenes_test("Chao1", psq, dataset)
print("Levene's Test Results for Chao1 Richness:")
print(levene_results_chao1)

# Perform Levene's test for Shannon diversity
levene_results_shannon <- perform_levenes_test("Shannon", psq, dataset)
print("Levene's Test Results for Shannon Diversity:")
print(levene_results_shannon)

# Perform Levene's test for Pielou evenness
levene_results_pielou <- perform_levenes_test("Pielou", psq, dataset)
print("Levene's Test Results for Pielou Evenness:")
print(levene_results_pielou)
```

```{r, alpha diversity plots}
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_alpha$new(dataset = dataset, group = "Region")
t1$cal_diff(method = "KW_dunn"); head(t1$res_diff)

# y_increase can adjust the distance from the letters to the highest point
A1 <- t1$plot_alpha(measure = "Chao1",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.85, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 8, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + labs(y = "Chao1 Richness")
                    axis.text.x = element_blank()) + labs(y = "Chao1 Richness")

A2 <- t1$plot_alpha(measure = "Shannon",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.85, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 8, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + labs(y = "Shannon Diversity")
                    axis.text.x = element_blank()) + labs(y = "Shannon Diversity")

t1 <- trans_alpha$new(dataset = dataset, group = "Region")
t1$cal_diff(method = "anova"); head(t1$res_diff)

A3 <- t1$plot_alpha(measure = "Pielou",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.85, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 8, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) + labs(y = "Pielou Evenness")
                    axis.text.x = element_blank()) + labs(y = "Pielou Evenness")

A1 + A2 + A3
```

```{r, Linear regression analysis}
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_alpha$new(dataset = dataset, group = "Region")

### Linear regression analysis

distance <- dataset$sample_table$Distance
alpha <- dataset$alpha_diversity
alpha$Distance <- distance

# Fit the linear regression model
model.cha <- lm(Chao1 ~ Distance, data = alpha)
model.sha <- lm(Shannon ~ Distance, data = alpha)
model.pie <- lm(Pielou ~ Distance, data = alpha)

# Summarize the model
summary(model.cha)
summary(model.sha)
summary(model.pie)

# Create the plot
L1 <- ggplot(alpha, aes(x = Distance, y = Chao1)) +
  geom_point(color = "gray50", alpha = 0.6) +  # Scatter plot of the data
  geom_smooth(method = "lm", color = "black", se = TRUE) +  # Add regression line with confidence interval
  labs(x = "km distance", y = "Chao1 Richness") +
  theme_bw() + theme(panel.grid = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1))

L2 <- ggplot(alpha, aes(x = Distance, y = Shannon)) +
  geom_point(color = "gray50", alpha = 0.6) +  # Scatter plot of the data
  geom_smooth(method = "lm", color = "black", se = TRUE) +  # Add regression line with confidence interval
  labs(x = "km distance", y = "Shannon Diversity") +
  theme_bw() + theme(panel.grid = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1))

L3 <- ggplot(alpha, aes(x = Distance, y = Pielou)) +
  geom_point(color = "gray50", alpha = 0.6) +  # Scatter plot of the data
  geom_smooth(method = "lm", color = "black", se = TRUE) +  # Add regression line with confidence interval
  labs(x = "km distance", y = "Pielou Evenness") +
  theme_bw() + theme(panel.grid = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1))

( L1 + L2 + L3 ) / ( A1 + A2 + A3 ) #L1 / L2 / L3
```

```{r, multiple comparison tests}
region <- dataset$sample_table$Region
alpha <- dataset$alpha_diversity
alpha$Region <- region

# Install and load the dunn.test package
if (!require(dunn.test)) {
  install.packages("dunn.test")
}
library(dunn.test)

# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(Chao1 ~ Region, data = alpha)
print(kruskal_result)

# Perform Dunn's test
dunn_result <- dunn.test(alpha$Chao1, alpha$Region, method = "bonferroni")

# Perform Dunn's test with Bonferroni adjustment for multiple comparisons
print(dunn_result)

# Perform Kruskal-Wallis test
kruskal_result <- kruskal.test(Shannon ~ Region, data = alpha)
print(kruskal_result)

# Perform Dunn's test with Bonferroni adjustment for multiple comparisons
dunn_result <- dunn.test(alpha$Shannon, alpha$Region, method = "bonferroni")

# Print the Dunn's test result
print(dunn_result)

# Perform ANOVA
anova_result <- aov(Pielou ~ Region, data = alpha)
summary(anova_result)

# Perform Duncan's test
duncan_result <- duncan.test(anova_result, "Region", console = TRUE)

# Print the Duncan's test result
print(duncan_result)
```

```{r, mOTU and other genes' alpha diversity plots}
dataset <- clone(otu.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_alpha$new(dataset = dataset, group = "Region")
t1$cal_diff(method = "anova"); head(t1$res_diff)

A4 <- t1$plot_alpha(measure = "Shannon",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Shannon Diversity")

A5 <- t1$plot_alpha(measure = "Pielou",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Pielou Evenness")
A4 + A5

dataset <- clone(otu.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_alpha$new(dataset = dataset, group = "Region")

### Linear regression analysis

distance <- dataset$sample_table$Distance
alpha <- dataset$alpha_diversity
alpha$Distance <- distance

# Fit the linear regression model
model.sha <- lm(Shannon ~ Distance, data = alpha)
model.pie <- lm(Pielou ~ Distance, data = alpha)

# Summarize the model
summary(model.sha)
summary(model.pie)

L4 <- ggplot(alpha, aes(x = Distance, y = Shannon)) +
  geom_point(color = "gray50", alpha = 0.6) +  # Scatter plot of the data
  geom_smooth(method = "lm", color = "black", se = TRUE) +  # Add regression line with confidence interval
  labs(x = "km distance", y = "Shannon Diversity") +
  theme_bw() + theme(panel.grid = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1))

L5 <- ggplot(alpha, aes(x = Distance, y = Pielou)) +
  geom_point(color = "gray50", alpha = 0.6) +  # Scatter plot of the data
  geom_smooth(method = "lm", color = "black", se = TRUE) +  # Add regression line with confidence interval
  labs(x = "km distance", y = "Pielou Evenness") +
  theme_bw() + theme(panel.grid = element_blank(), axis.text.y = element_text(angle = 90, hjust = 1))

dataset <- clone(mrg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_alpha$new(dataset = dataset, group = "Region")
t1$cal_diff(method = "anova"); head(t1$res_diff)

# y_increase can adjust the distance from the letters to the highest point
A6 <- t1$plot_alpha(measure = "Chao1",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Chao1 Richness")

A7 <- t1$plot_alpha(measure = "Shannon",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Shannon Diversity")

A8 <- t1$plot_alpha(measure = "Pielou",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Pielou Evenness")

A6 + A7 + A8

dataset <- clone(mge.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_alpha$new(dataset = dataset, group = "Region")
t1$cal_diff(method = "anova"); head(t1$res_diff)

# y_increase can adjust the distance from the letters to the highest point
A9 <- t1$plot_alpha(measure = "Chao1",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Chao1 Richness")

A10 <- t1$plot_alpha(measure = "Shannon",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Shannon Diversity")

A11 <- t1$plot_alpha(measure = "Pielou",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Pielou Evenness")

A9 + A10 + A11

dataset <- clone(vfg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

t1 <- trans_alpha$new(dataset = dataset, group = "Region")
t1$cal_diff(method = "anova"); head(t1$res_diff)

# y_increase can adjust the distance from the letters to the highest point
A12 <- t1$plot_alpha(measure = "Chao1",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Chao1 Richness")

A13 <- t1$plot_alpha(measure = "Shannon",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Shannon Diversity")

A14 <- t1$plot_alpha(measure = "Pielou",
              add_sig_text_size = 4,
              #boxplot_add = "dotplot",
              order_x_mean = FALSE,
              #shape = "Group",
              #add_sample_label = "Group",
              #shape_values = c(17,2),
              color_values = brewer.pal(n = 8, name = "Dark2")) + #scale_color_grey() +
              geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 0.5, outlier.shape = NA) + theme_bw() +
              theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
                    #axis.text.x = element_text(size = 10, angle = 0))
                    axis.text.x = element_blank()) + labs(y = "Pielou Evenness")

A12 + A13 + A14

(A4 + A5) / (L4 + L5) / (A6 + A7 + A8) / (A9 + A10 + A11) / (A12 + A13 + A14) + patchwork::plot_annotation(tag_levels = 'A') + theme(plot.tag = element_text(face = "bold"))
```

### Step 3C: Statistical Analysis - Beta diversity estimation 

```{r}
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Subtype"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment
#dataset$cal_betadiv(unifrac = FALSE); class(dataset$beta_diversity) ## Create trans_beta object and measure parameter - dataset$beta_diversity

# measure parameter must be one of names(dataset$beta_diversity)
t1 <- trans_beta$new(dataset = dataset, group = "Region", measure = "bray")
# PCoA, PCA and NMDS are available
t1$cal_ordination(method = "PCoA"); class(t1$res_ordination) # t1$res_ordination is the ordination result list
# plot the PCoA result with confidence ellipse
B1 <- t1$plot_ordination(plot_color = "Region", 
                         plot_shape = "Region", plot_type = c("chull"), ellipse_chull_alpha = 0,
                         color_values = RColorBrewer::brewer.pal(8, "Dark2"), add_sample_label = "Sample") +
                         theme_bw() + theme(panel.grid = element_blank(), legend.position = "none", axis.text.y = element_text(angle = 90, hjust = 0.5)) +
                         ggalt::geom_encircle(aes(s_shape = 1, expand = 0, group = Region, fill = Region, colour = Region), alpha = 0.5) +
                         annotate("text", x = -0.15, y = -0.3, label = "Bothnian Bay", colour = "#1B9E77") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Bothnian Sea", colour = "#D95F02") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Stockholm", colour = "#7570B3") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Sörmland", colour = "#E7298A") +
                         #annotate("text", x = -0.15, y = -0.08, label = "Dead Zone (Sörmland)", colour = "#66A61E") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Östergötland", colour = "#E6AB02") +
                         #annotate("text", x = -0.2, y = 0, label = "Dead Zone (Mid-South)", colour = "#A6761D") +
                         annotate("text", x = -0.3, y = 0.1, label = "Southern Baltic", colour = "#666666")
B1

# manova for specified group set: such as "Region"
t1$cal_manova(); t1$res_manova

t1$cal_anosim(group = "Region"); t1$res_anosim
t1$cal_betadisper(); t1$res_betadisper

# calculate and plot sample distances within groups
t1$cal_group_distance(within_group = TRUE)
# return t1$res_group_distance
# perform ANOVA
t1$cal_group_distance_diff(method = "anova"); t1$res_group_distance_diff
# plot_group_order parameter can be used to adjust orders in x axis
P1 <- t1$plot_group_distance(boxplot_add = "mean",
                             plot_group_order = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland",
                                                  "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic")) +
      geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 1, outlier.shape = NA) + theme_bw() +
      #geom_smooth(method = loess, se = TRUE, aes(group = 1)) +
      theme(legend.position = "none", axis.text.y = element_text(size = 10, angle = 90, hjust = 0.5) 
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.text.x = element_text(size = 10, angle = 40, hjust = 1))
P1
```

```{r}
dataset <- clone(otu.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Subtype"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment
#dataset$cal_betadiv(unifrac = FALSE); class(dataset$beta_diversity) ## Create trans_beta object and measure parameter - dataset$beta_diversity

# measure parameter must be one of names(dataset$beta_diversity)
t1 <- trans_beta$new(dataset = dataset, group = "Region", measure = "bray")
# PCoA, PCA and NMDS are available
t1$cal_ordination(method = "PCoA"); class(t1$res_ordination) # t1$res_ordination is the ordination result list
# plot the PCoA result with confidence ellipse
B2 <- t1$plot_ordination(plot_color = "Region", 
                         plot_shape = "Region", plot_type = c("chull"), ellipse_chull_alpha = 0,
                         color_values = RColorBrewer::brewer.pal(8, "Dark2")) +
                         theme_bw() + theme(panel.grid = element_blank(), legend.position = "none", axis.text.y = element_text(angle = 90, hjust = 0.5)) +
                         ggalt::geom_encircle(aes(s_shape = 1, expand = 0, group = Region, fill = Region, colour = Region), alpha = 0.5)
B2

# manova for specified group set: such as "Region"
t1$cal_manova(); t1$res_manova

#t1$cal_anosim(group = "Region"); t1$res_anosim
t1$cal_betadisper(); t1$res_betadisper

dataset <- clone(mrg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Subtype"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment
#dataset$cal_betadiv(unifrac = FALSE); class(dataset$beta_diversity) ## Create trans_beta object and measure parameter - dataset$beta_diversity

# measure parameter must be one of names(dataset$beta_diversity)
t1 <- trans_beta$new(dataset = dataset, group = "Region", measure = "bray")
# PCoA, PCA and NMDS are available
t1$cal_ordination(method = "PCoA"); class(t1$res_ordination) # t1$res_ordination is the ordination result list
# plot the PCoA result with confidence ellipse
B3 <- t1$plot_ordination(plot_color = "Region", 
                         plot_shape = "Region", plot_type = c("chull"), ellipse_chull_alpha = 0,
                         color_values = RColorBrewer::brewer.pal(8, "Dark2")) +
                         theme_bw() + theme(panel.grid = element_blank(), legend.position = "none", axis.text.y = element_text(angle = 90, hjust = 0.5)) +
                         ggalt::geom_encircle(aes(s_shape = 1, expand = 0, group = Region, fill = Region, colour = Region), alpha = 0.5)
B3

# manova for specified group set: such as "Region"
t1$cal_manova(); t1$res_manova

#t1$cal_anosim(group = "Region"); t1$res_anosim
t1$cal_betadisper(); t1$res_betadisper

dataset <- clone(mge.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Subtype"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment
#dataset$cal_betadiv(unifrac = FALSE); class(dataset$beta_diversity) ## Create trans_beta object and measure parameter - dataset$beta_diversity

# measure parameter must be one of names(dataset$beta_diversity)
t1 <- trans_beta$new(dataset = dataset, group = "Region", measure = "bray")
# PCoA, PCA and NMDS are available
t1$cal_ordination(method = "PCoA"); class(t1$res_ordination) # t1$res_ordination is the ordination result list
# plot the PCoA result with confidence ellipse
B4 <- t1$plot_ordination(plot_color = "Region", 
                         plot_shape = "Region", plot_type = c("chull"), ellipse_chull_alpha = 0,
                         color_values = RColorBrewer::brewer.pal(8, "Dark2")) +
                         theme_bw() + theme(panel.grid = element_blank(), legend.position = "none", axis.text.y = element_text(angle = 90, hjust = 0.5)) +
                         ggalt::geom_encircle(aes(s_shape = 1, expand = 0, group = Region, fill = Region, colour = Region), alpha = 0.5)
B4

# manova for specified group set: such as "Region"
t1$cal_manova(); t1$res_manova

#t1$cal_anosim(group = "Region"); t1$res_anosim
t1$cal_betadisper(); t1$res_betadisper

dataset <- clone(vfg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Subtype"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment
#dataset$cal_betadiv(unifrac = FALSE); class(dataset$beta_diversity) ## Create trans_beta object and measure parameter - dataset$beta_diversity

# measure parameter must be one of names(dataset$beta_diversity)
t1 <- trans_beta$new(dataset = dataset, group = "Region", measure = "bray")
# PCoA, PCA and NMDS are available
t1$cal_ordination(method = "PCoA"); class(t1$res_ordination) # t1$res_ordination is the ordination result list
# plot the PCoA result with confidence ellipse
B5 <- t1$plot_ordination(plot_color = "Region", 
                         plot_shape = "Region", plot_type = c("chull"), ellipse_chull_alpha = 0,
                         color_values = RColorBrewer::brewer.pal(8, "Dark2")) +
                         theme_bw() + theme(panel.grid = element_blank(), legend.position = "none", axis.text.y = element_text(angle = 90, hjust = 0.5)) +
                         ggalt::geom_encircle(aes(s_shape = 1, expand = 0, group = Region, fill = Region, colour = Region), alpha = 0.5)
B5

# manova for specified group set: such as "Region"
t1$cal_manova(); t1$res_manova

#t1$cal_anosim(group = "Region"); t1$res_anosim
t1$cal_betadisper(); t1$res_betadisper

gridExtra::grid.arrange(B2, B3, B4, B5, nrow = 1)
```

### Step 3D: Statistical Analysis - Partial least squares path modeling (PLS-PM) analysis 

```{r}
# Load necessary library
library("plspm")
library("turner")
library("microbiome")
library("metagMisc")

set.seed(123456) # for reproducibility, remove for normal use

# Create dataframe with datasets that only have genes that are found in at least 5% of samples.

#Based on UN-aggregated datasets
baltic.df <- cbind(t(as.data.frame(abundances(phyloseq_filter_prevalence(otu.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   #t(as.data.frame(abundances(phyloseq_filter_prevalence(transform_sample_counts(otu.psq, function(x) ifelse(x > 0, 1, 0)), prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(arg.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(mge.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(vfg.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(mrg.psq, prev.trh = 0.05, abund.trh = NULL)))));
dim(baltic.df)

# define path matrix (inner model)
Microbes <- c(0,0,0,0,0)
MGE <- c(1,0,0,0,0)
VFG <- c(1,1,0,0,0)
ARG <- c(1,1,1,0,0)
MRG <- c(1,1,1,1,0)

baltic.path <- rbind(Microbes, MGE, VFG, ARG, MRG)

# define list of blocks (outer model)
baltic.blocks <- list(1:2319, 3579:9947, 9948:10715, 2320:3578, 10716:11928) # using UN-aggregated dataset.

# vector of modes (reflective indicators)
baltic.modes <- rep("A", 5) 

# apply plspm with bootstrap validation
baltic.pls <- plspm(baltic.df, baltic.path, baltic.blocks, modes = baltic.modes, boot.val = 1000)

# default print
print(baltic.pls)

# summary of results
summary(baltic.pls)

baltic.pls$gof

# plot inner model results
plot(baltic.pls, what = "inner",
     colpos = "#4169E1", colneg = "#F08080", arr.width = 0.3,
     box.prop = 0.55, box.size = 0.08, box.cex = 1, box.col = c("#E78AC3","#8DA0CB","#FC8D62","#E5C494","#A6D854"), lcol = "black",
     txt.col = "black", arr.pos = 0.5, cex.txt = 0.9)

# plot outer model loadings
#plot(baltic.pls, what = "loadings")

# plot outer model weights
#plot(baltic.pls, what = "weights")
```

```{r}
# Load necessary libraries
library(ggplot2)
library(reshape2)

# Create the data frame
data <- data.frame(
  Relationships = c("Microbes - MGE", "Microbes - VFG", "Microbes - ARG", "Microbes - MRG",
                    "MGE - VFG", "MGE - ARG", "MGE - MRG", "VFG - ARG", "VFG - MRG"),
  Direct = c(0.9680, 0.4604, 0.6016, -0.3293, -1.3748, 0.4387, -0.5820, 0.0483, 0.0939),
  Indirect = c(0, -1.3308, 0.3826, -0.6451, 0, -0.0664, -0.1291, 0, 0),
  Total = c(0.9680, -0.8704, 0.9842, -0.9744, -1.3748, 0.3723, -0.7111, 0.0483, 0.0939)
)

# Melt the data frame for ggplot2
melted_data <- melt(data, id.vars = "Relationships")

# Plot using ggplot2
ggplot(melted_data, aes(x = Relationships, y = value, fill = variable)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(#title = "Direct, Indirect, and Total Effects",
       x = "Relationships",
       y = "Coefficients") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1), panel.grid = element_blank(), legend.position = "top") +
  scale_fill_manual(values = c("Direct" = "#D95F02", "Indirect" = "#7570B3", "Total" = "#666666"))
```

```{r eval = FALSE, NOT PERFORMED}
# Load necessary library
library("plspm")
library("turner")
library("microbiome")
library("metagMisc")

set.seed(123456) # for reproducibility, remove for normal use

# Create dataframe with datasets that only have genes that are found in at least 5% of samples.

#Based on UN-aggregated datasets
baltic.df <- cbind(t(as.data.frame(abundances(phyloseq_filter_prevalence(otu.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(arg.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(mge.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(vfg.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   t(as.data.frame(abundances(phyloseq_filter_prevalence(mrg.psq, prev.trh = 0.05, abund.trh = NULL)))),
                   metadata.log[, c(7:10,13:15)]
                   );


dim(baltic.df)

# define path matrix (inner model)
ENV <- c(0,0,0,0,0,0)
Microbes <- c(1,0,0,0,0,0)
MGE <- c(1,1,0,0,0,0)
VFG <- c(1,1,1,0,0,0)
ARG <- c(1,1,1,1,0,0)
MRG <- c(1,1,1,1,0,0)

baltic.path <- rbind(ENV, Microbes, MGE, VFG, ARG, MRG)

# define list of blocks (outer model)
baltic.blocks <- list(11929:11935, 1:2319, 3579:9947, 9948:10715, 2320:3578, 10716:11928) # using UN-aggregated dataset.

# vector of modes (reflective indicators)
baltic.modes <- rep("A", 6) 

# apply plspm with bootstrap validation
baltic.pls <- plspm(baltic.df, baltic.path, baltic.blocks, modes = baltic.modes, boot.val = 1000)

# default print
print(baltic.pls)

# summary of results
summary(baltic.pls)

# plot inner model results
plot(baltic.pls, what = "inner",
     colpos = "#4169E1", colneg = "#F08080", arr.width = 0.3,
     box.prop = 0.55, box.size = 0.08, box.cex = 1, box.col = c("#E78AC3","#8DA0CB","#FC8D62","#E5C494","#A6D854","#1B9E77"), lcol = "black",
     txt.col = "black", arr.pos = 0.5, cex.txt = 0.9)

# plot outer model loadings
#plot(baltic.pls, what = "loadings")

# plot outer model weights
#plot(baltic.pls, what = "weights")
```

### Step 3E: Statistical Analysis - Procrustes test between gene datasets 

```{r}
library("vegan")

# Extract beta diversity matrices (Bray-Curtis) for each dataset
arg.bc <- arg.mco$beta_diversity$bray
otu.bc <- otu.mco$beta_diversity$bray
mrg.bc <- mrg.mco$beta_diversity$bray
mge.bc <- mge.mco$beta_diversity$bray
vfg.bc <- vfg.mco$beta_diversity$bray

# Hellinger transformation for PCA
arg.bc.hel <- decostand(arg.bc, method = "hellinger")
arg.bc.hel.pca <- rda(arg.bc.hel)

otu.bc.hel <- decostand(otu.bc, method = "hellinger")
otu.bc.hel.pca <- rda(otu.bc.hel)

mrg.bc.hel <- decostand(mrg.bc, method = "hellinger")
mrg.bc.hel.pca <- rda(mrg.bc.hel)

mge.bc.hel <- decostand(mge.bc, method = "hellinger")
mge.bc.hel.pca <- rda(mge.bc.hel)

vfg.bc.hel <- decostand(vfg.bc, method = "hellinger")
vfg.bc.hel.pca <- rda(vfg.bc.hel)

# Test between ARG and OTU
arg.otu.pro <- procrustes(X = arg.bc.hel.pca, Y = otu.bc.hel.pca, symmetric = TRUE); summary(arg.otu.pro)
arg.otu.prt <- protest(X = arg.bc.hel.pca, Y = otu.bc.hel.pca, scores = "sites", permutations = 999); arg.otu.prt

arg.mrg.pro <- procrustes(X = arg.bc.hel.pca, Y = mrg.bc.hel.pca, symmetric = TRUE); summary(arg.mrg.pro)
arg.mrg.prt <- protest(X = arg.bc.hel.pca, Y = mrg.bc.hel.pca, scores = "sites", permutations = 999); arg.mrg.prt

arg.mge.pro <- procrustes(X = arg.bc.hel.pca, Y = mge.bc.hel.pca, symmetric = TRUE); summary(arg.mge.pro)
arg.mge.prt <- protest(X = arg.bc.hel.pca, Y = mge.bc.hel.pca, scores = "sites", permutations = 999); arg.mge.prt

arg.vfg.pro <- procrustes(X = arg.bc.hel.pca, Y = vfg.bc.hel.pca, symmetric = TRUE); summary(arg.vfg.pro)
arg.vfg.prt <- protest(X = arg.bc.hel.pca, Y = vfg.bc.hel.pca, scores = "sites", permutations = 999); arg.vfg.prt

#plot(arg.otu.pro, kind = 1)
plot(arg.otu.pro, kind = 1, main = "ARGs and Microbes", ar.col = "gray")
lines(arg.otu.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.otu.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.otu.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#A6D854")

#plot(arg.mrg.pro, kind = 1)
plot(arg.mrg.pro, kind = 1, main = "ARGs and MRGs", ar.col = "gray")
lines(arg.mrg.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.mrg.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.mrg.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#E78AC3")

#plot(arg.mge.pro, kind = 1)
plot(arg.mge.pro, kind = 1, main = "ARGs and MGEs", ar.col = "gray")
lines(arg.mge.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.mge.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.mge.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#E5C494")

#plot(arg.vfg.pro, kind = 1)
plot(arg.vfg.pro, kind = 1, main = "ARGs and VFGs", ar.col = "gray")
lines(arg.vfg.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.vfg.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.vfg.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FC8D62")
```

### Step 3F: Statistical Analysis - Mantel statistic based on Pearson's product-moment correlation 

```{r}
library("vegan")

arg.bc <- arg.mco$beta_diversity$bray
otu.bc <- otu.mco$beta_diversity$bray
mrg.bc <- mrg.mco$beta_diversity$bray
mge.bc <- mge.mco$beta_diversity$bray
vfg.bc <- vfg.mco$beta_diversity$bray

mantel(arg.bc, otu.bc, method = "pearson", permutations = 999)
mantel(arg.bc, mrg.bc, method = "pearson", permutations = 999)
mantel(arg.bc, mge.bc, method = "pearson", permutations = 999)
mantel(arg.bc, vfg.bc, method = "pearson", permutations = 999)
```

### Step 3G: Statistical Analysis - Distance-based redundancy analysis (dbRDA) 

```{r}
# Clone the dataset and set Region as a factor with specified levels
dataset <- clone(arg.mco)
dataset$sample_table$Region <- factor(dataset$sample_table$Region, levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", 
                                                                             "Dead Zone (Sörmland)", "Östergötland", 
                                                                             "Dead Zone (Mid-South)", "Southern Baltic"))

# Rename columns with special characters
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Depth")] <- "Depth***"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Temperature")] <- "Temperature***"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Salinity")] <- "Salinity***"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "O2")] <- "Dissolved O2**"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "C.N.Ratio")] <- "C:N Ratio***"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Delta.15N")] <- "δ15N"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Delta.13C")] <- "δ13C***"

# Initialize trans_env object and add environmental data
t1 <- trans_env$new(dataset = dataset, add_data = dataset$sample_table[, c(7:10, 13:15)])

# Perform dbRDA analysis using Bray-Curtis distance
t1$cal_ordination(method = "dbRDA", use_measure = "bray")

# Transform ordination results for visualization
t1$trans_ordination(adjust_arrow_length = TRUE, max_perc_env = 1.5)

# Plot ordination with environmental variables and color by Region
t1$plot_ordination(plot_color = "Region", plot_type = "point",
                   color_values = RColorBrewer::brewer.pal(8, "Dark2")) +
  theme_bw() +
  theme(panel.grid = element_blank(), legend.position = "right")

# View ordination terms
t1$res_ordination_terms

# Perform ANOVA on ordination axes
t1$cal_ordination_anova()
t1$res_ordination_axis

# Perform envfit analysis
t1$cal_ordination_envfit()
t1$res_ordination_envfit


t1$res_ordination_R2
```

### Step 3H: Statistical Analysis - Distance-decay relationships of Bray-Curtis dissimilarity and haversine distance

```{r}
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

# Remove the dead zones.
#dataset$sample_table %<>% subset(Region == "Bothnian Bay" | Region == "Bothnian Sea" | Region == "Stockholm" | Region == "Sörmland" | Region == "Östergötland" | Region == "Southern Baltic"); dataset$tidy_dataset(); dataset ## Remove groups

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D1 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", line_alpha = 0.25, line_se_color = "#8DA0CB", lm_equation = FALSE, line_color = "#8DA0CB",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + theme(panel.grid = element_blank()) + geom_point(colour = "#8DA0CB", alpha = 0.05) + ggtitle("ARGs")
D1

dataset <- clone(otu.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
# Remove the dead zones.
#dataset$sample_table %<>% subset(Region == "Bothnian Bay" | Region == "Bothnian Sea" | Region == "Stockholm" | Region == "Sörmland" | Region == "Östergötland" | Region == "Southern Baltic"); dataset$tidy_dataset(); dataset ## Remove groups

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D2 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", line_alpha = 0.25, line_se_color = "#A6D854", lm_equation = FALSE, line_color = "#A6D854",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + theme(panel.grid = element_blank()) + geom_point(colour = "#A6D854", alpha = 0.05) + ggtitle("Microbes")
D2

dataset <- clone(mrg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
# Remove the dead zones.
#dataset$sample_table %<>% subset(Region == "Bothnian Bay" | Region == "Bothnian Sea" | Region == "Stockholm" | Region == "Sörmland" | Region == "Östergötland" | Region == "Southern Baltic"); dataset$tidy_dataset(); dataset ## Remove groups

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D3 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", line_alpha = 0.25, line_se_color = "#E78AC3", lm_equation = FALSE, line_color = "#E78AC3",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + theme(panel.grid = element_blank()) + geom_point(colour = "#E78AC3", alpha = 0.05) + ggtitle("MRGs")
D3

dataset <- clone(mge.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
# Remove the dead zones.
#dataset$sample_table %<>% subset(Region == "Bothnian Bay" | Region == "Bothnian Sea" | Region == "Stockholm" | Region == "Sörmland" | Region == "Östergötland" | Region == "Southern Baltic"); dataset$tidy_dataset(); dataset ## Remove groups

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D4 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", line_alpha = 0.25, line_se_color = "#E5C494", lm_equation = FALSE, line_color = "#E5C494",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + theme(panel.grid = element_blank()) + geom_point(colour = "#E5C494", alpha = 0.05) + ggtitle("MGEs")
D4

dataset <- clone(vfg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
# Remove the dead zones.
#dataset$sample_table %<>% subset(Region == "Bothnian Bay" | Region == "Bothnian Sea" | Region == "Stockholm" | Region == "Sörmland" | Region == "Östergötland" | Region == "Southern Baltic"); dataset$tidy_dataset(); dataset ## Remove groups

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D5 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", line_alpha = 0.25, line_se_color = "#FC8D62", lm_equation = FALSE, line_color = "#FC8D62",
                         x_axis_title = "km distance", y_axis_title = "VFG Bray-Curtis distance") + theme_bw() + theme(panel.grid = element_blank()) + geom_point(colour = "#FC8D62", alpha = 0.05) + ggtitle("VFGs")
D5

#D1 + (D2 + D3) / (D4 + D5)
ggpubr::ggarrange(D1, D2, D3, D4, D5, nrow = 1)
```

### Step 3I: Statistical Analysis - Pearson correlation of environmnetal variables and ARG diversity and composition 

```{r}
# Clone the dataset and set factor levels for the 'Region' column
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)", "Southern Baltic"))

# Create an object to assess the relationship between environmental factors and alpha diversity
# Use 'env_cols' to specify the columns containing environmental data
t1 <- trans_env$new(dataset = dataset, env_cols = c(7:10, 13:15), complete_na = TRUE)

# Calculate the correlation using specified alpha diversity metrics
t1$cal_cor(add_abund_table = dataset$alpha_diversity[c("Chao1", "Shannon", "Pielou")], p_adjust_type = "Env", cor_method = "pearson", p_adjust_method = "fdr")

# Plot the correlation results
H1 <- t1$plot_cor(cluster_ggplot = "both", color_vector = c("#F08080", "white", "#4169E1"))
t1$res_cor

# Plot scatter fits for various environmental factors against Bray-Curtis distance

# Salinity
E1 <- t1$plot_scatterfit(
  x = "Salinity",
  y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)],
  type = "cor",
  point_size = 2,
  point_alpha = 0.01,
  label.x.npc = "left",
  label.y.npc = "bottom",
  line_alpha = 0.25,
  line_se_color = "gray70",
  x_axis_title = "Salinity",
  y_axis_title = "Bray-Curtis distance"
) + theme_bw() + theme(panel.grid = element_blank())

# Temperature
E2 <- t1$plot_scatterfit(
  x = "Temperature",
  y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)],
  type = "cor",
  point_size = 2,
  point_alpha = 0.01,
  label.x.npc = "left",
  label.y.npc = "bottom",
  line_alpha = 0.25,
  line_se_color = "gray70",
  x_axis_title = "Temperature (°C)",
  y_axis_title = "Bray-Curtis distance"
) + theme_bw() + theme(panel.grid = element_blank())

# Depth
E3 <- t1$plot_scatterfit(
  x = "Depth",
  y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)],
  type = "cor",
  point_size = 2,
  point_alpha = 0.01,
  label.x.npc = "left",
  label.y.npc = "bottom",
  line_alpha = 0.25,
  line_se_color = "gray70",
  x_axis_title = "Depth (m)",
  y_axis_title = "Bray-Curtis distance"
) + theme_bw() + theme(panel.grid = element_blank())

# Dissolved O2
E4 <- t1$plot_scatterfit(
  x = "O2",
  y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)],
  type = "cor",
  point_size = 2,
  point_alpha = 0.01,
  label.x.npc = "left",
  label.y.npc = "bottom",
  line_alpha = 0.25,
  line_se_color = "gray70",
  x_axis_title = expression(Dissolved~O[2]),
  y_axis_title = "Bray-Curtis distance"
) + theme_bw() + theme(panel.grid = element_blank())

# C:N Ratio
E5 <- t1$plot_scatterfit(
  x = "C.N.Ratio",
  y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)],
  type = "cor",
  point_size = 2,
  point_alpha = 0.01,
  label.x.npc = "left",
  label.y.npc = "bottom",
  line_alpha = 0.25,
  line_se_color = "gray70",
  x_axis_title = "C:N Ratio",
  y_axis_title = "Bray-Curtis distance"
) + theme_bw() + theme(panel.grid = element_blank())

# δ15N
E6 <- t1$plot_scatterfit(
  x = "Delta.15N",
  y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)],
  type = "cor",
  point_size = 2,
  point_alpha = 0.01,
  label.x.npc = "left",
  label.y.npc = "bottom",
  line_alpha = 0.25,
  line_se_color = "gray70",
  x_axis_title = expression(delta^15*N~'\u2030'),
  y_axis_title = "Bray-Curtis distance"
) + theme_bw() + theme(panel.grid = element_blank())

# δ13C
E7 <- t1$plot_scatterfit(
  x = "Delta.13C",
  y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)],
  type = "cor",
  point_size = 2,
  point_alpha = 0.01,
  label.x.npc = "left",
  label.y.npc = "bottom",
  line_alpha = 0.25,
  line_se_color = "gray70",
  x_axis_title = expression(delta^13*C~'\u2030'),
  y_axis_title = "Bray-Curtis distance"
) + theme_bw() + theme(panel.grid = element_blank())

# Combine the plots into a single display
#( E1 | E2 | E3 | E4 ) / ( patchwork::plot_spacer() | E5 | E6 | E7 )
( E1  | E2 | E3 ) / ( E4  | E5 | E6 ) / ( patchwork::plot_spacer() | patchwork::plot_spacer() | E7 )
```

### Step 3J: Statistical Analysis - Generalized additive models (GAMs) of the environmental factors and alpha diversity 

```{r, multiple regression analysis}
# Load required library for GAMs
library(mgcv)

# Assuming 'arg.mco' is your original dataset

# Clone the dataset and factorize Region column
dataset <- clone(arg.mco)
dataset$sample_table$Region <- factor(dataset$sample_table$Region,
                                      levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", 
                                                 "Sörmland", "Dead Zone (Sörmland)", 
                                                 "Östergötland", "Dead Zone (Mid-South)", 
                                                 "Southern Baltic"))

# Initialize trans_env object to add environmental data
env <- dataset$sample_table
alpha <- dataset$alpha_diversity

# Merge environmental and alpha diversity data
t2 <- merge(env, alpha, by = "row.names", all = TRUE)
rownames(t2) <- t2[, 1]
t2 <- t2[, -1]

# Define GAM formula
formula <- Chao1 ~ s(Depth) + s(Salinity) + s(Temperature) + s(O2) + s(C.N.Ratio) + s(Delta.15N) + s(Delta.13C)

# Fit GAM for Chao1
model.chao <- gam(formula, data = t2)
summary(model.chao)

# Fit GAM for Shannon
formula <- Shannon ~ s(Depth) + s(Salinity) + s(Temperature) + s(O2) + s(C.N.Ratio) + s(Delta.15N) + s(Delta.13C)
model.shannon <- gam(formula, data = t2)
summary(model.shannon)

# Fit GAM for Pielou
formula <- Pielou ~ s(Depth) + s(Salinity) + s(Temperature) + s(O2) + s(C.N.Ratio) + s(Delta.15N) + s(Delta.13C)
model.pielou <- gam(formula, data = t2)
summary(model.pielou)
```

```{r}
# Install and load required packages (if not already installed)

library(ggplot2)
library(mgcv)
library(visreg)

# Assuming model.chao, model.shannon, model.pielou are already fitted GAM models

# Function to plot smooth terms
plot_smooth_terms <- function(model, response_name) {
  visreg(model, scale = "response", ylab = response_name) +
    ggtitle(paste("Smooth Terms for", response_name))
}

# Plot smooth terms for Chao1
plot_smooth_terms(model.chao, "Chao1")

# Plot smooth terms for Shannon
plot_smooth_terms(model.shannon, "Shannon")

# Plot smooth terms for Pielou
plot_smooth_terms(model.pielou, "Pielou")
```

### Step 3K: Statistical Analysis - Random Forest to identify important environmental variables influence on ARG composition

```{r}
# Install and load necessary packages
packages <- c("Boruta", "parallel", "rsample", "randomForest", "caret", "gridExtra", "multiROC", "rfPermute")

for (pkg in packages) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
    library(pkg, character.only = TRUE)
  }
}

# Load required libraries
library(microeco)
library(ggplot2)  # For plotting (optional)

# Clone the dataset and factorize Region column
dataset <- clone(arg.mco)
dataset$sample_table$Region <- factor(dataset$sample_table$Region,
                                      levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", 
                                                 "Sörmland", "Dead Zone (Sörmland)", 
                                                 "Östergötland", "Dead Zone (Mid-South)", 
                                                 "Southern Baltic"))

# Initialize trans_env object to add environmental data
t1 <- trans_env$new(dataset = dataset, env_cols = c(7:10, 13:15), complete_na = TRUE)
tmp <- t1$data_env %>% t %>% as.data.frame

# Initialize trans_classifier object for machine learning
t2 <- trans_classifier$new(dataset = dataset, x.predictors = tmp, y.response = "Salinity")
t2$cal_preProcess(method = c("center", "scale", "nzv"))

# Train the Random Forest model
t2$cal_train(method = "rf")

# Generate feature importance scores with permutation-based significance
t2$cal_feature_imp(rf_feature_sig = TRUE, num.rep = 1000)

env_labs <- rev(c("Salinity", "Temperature", expression(Dissolved~O[2]), "C:N Ratio", expression(delta^13*C), "Water Depth", expression(delta^15*N)))

# Plot feature importance
R1 <- t2$plot_feature_imp(show_sig_group = TRUE, coord_flip = TRUE, width = 0.6, add_sig = TRUE, color_values = "gray25") + 
  theme_bw() + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 0, hjust = 1), panel.grid = element_blank()) + 
  ylab("Mean Decrease Accuracy") + scale_x_discrete(labels = env_labs)

print(R1)
```

### Step 3L: Statistical Analysis - Spearman correlation analysis between the environmental factors and ARG Types 

```{r}
library(dplyr)
library(tidyr)
library(ggplot2)

# Load psq file.
psq <- arg.psq

# Compute correlations with p values and store in a dataframe
correlations_df <- psq %>%
  tax_model(trans = "clr", rank = "Type", 
            variables = list("Depth", "Salinity", "Temperature", "O2", "C.N.Ratio", "Delta.15N", "Delta.13C"),
            type = microViz::cor_test, method = "spearman", return_psx = FALSE, verbose = FALSE) %>%
  tax_models2stats(rank = "Type")

# Calculate FDR adjusted p-values
correlations_df <- correlations_df %>%
  mutate(p.FDR = p.adjust(p.value, method = "fdr"))

# Hierarchical clustering of taxa based on correlation estimates
taxa_hclust <- correlations_df %>%
  select(term, taxon, estimate) %>%
  pivot_wider(id_cols = taxon, names_from = term, values_from = estimate) %>%
  column_to_rownames("taxon") %>%
  as.matrix() %>%
  dist(method = "euclidean") %>%
  hclust(method = "mcquitty")

taxa_order <- taxa_hclust$labels[taxa_hclust$order]
env_labs <- c("C:N Ratio", expression(delta^13*C), expression(delta^15*N), "Water Depth", expression(Dissolved~O[2]), "Salinity", "Temperature")

# Plotting
ggplot(correlations_df, aes(x = term, y = taxon)) +
  geom_raster(aes(fill = estimate)) +
  geom_point(data = filter(correlations_df, p.value < 0.05), shape = "asterisk") +
  geom_point(data = filter(correlations_df, p.FDR < 0.05), shape = "circle", size = 3) +
  scale_y_discrete(limits = taxa_order) +
  scale_fill_distiller(palette = "PuOr", direction = -1, limits = c(-0.7, 0.7)) +
  theme_bw() +
  theme(axis.text.x = element_text(face = "bold", angle = 35, hjust = 1, size = 10),
        legend.position = "left", axis.ticks.length.y = unit(10, "pt")) +
  labs(x = NULL, y = NULL, fill = "Spearman's Rank Correlation") +
  scale_x_discrete(labels = env_labs) #+
  #scale_y_dendrogram(guide = guide_dendro(position = "right"), hclust = taxa_hclust)

# Write correlations to CSV file
write.csv(correlations_df, "out/correlations.tsv", quote = FALSE)
```

```{r}
### [NOT PERFORMED] Pearson correlation of environmnetal variables and ARG types. 
library(microbiomeViz)

# Assuming 'arg.psq' is already defined and psq is your ARG data

psq %>%
  tax_agg("Type") %>%
  cor_heatmap(
    vars = c("Depth", "Salinity", "Temperature", "O2", "TN", "TC", "C.N.Ratio", "Delta.13C", "Delta.15N"),
    taxa = tax_top(psq, 50, by = max, rank = "Type"),
    tax_anno = taxAnnotation(
      Prev. = anno_tax_prev(ylim = 0:1),
      Log10. = anno_tax_box(trans = "log10", zero_replace = "halfmin")
    ),
    colors = heat_palette(palette = "Blue-Red 2", sym = TRUE, rev = TRUE),
    numbers = heat_numbers(decimals = 1, col = "white", fontface = "plain"),
    taxa_side = "right"
  )
```

### Others - Custom legend 

```{r}
# Define the colors and text for the legend
legend_colors <- c("#8DA0CB","#A6D854","#E78AC3","#E5C494","#FC8D62")
legend_text <- c("ARG", "Microbes", "MRG", "MGE", "VFG")

# Create a data frame for legend entries
legend_df <- data.frame(x = rep(1, length(legend_text)), y = seq(length(legend_text)), legend_text = legend_text, legend_colors = factor(legend_text, levels = legend_text))

# Create a ggplot with custom legend
custom_legend_plot <- ggplot(legend_df, aes(x = x, y = y, color = legend_colors)) + geom_point(size = 5) + scale_color_manual(values = legend_colors, guide = guide_legend(title = "Legend")) + theme_bw() + theme(legend.position = "right")

# Print the custom legend plot
print(custom_legend_plot)
```

### Step XX: Database Comparison: Load datasets

```{r}
library("microbiomer")

# Import counts annotation data from Excel file and define row names
amr.tbl <- read_excel("data/data.xlsx", sheet = "amrpp")
amr.tbl <- column_to_rownames(amr.tbl, loc = 1)

# Create phyloseq object
amr.psq <- phyloseq(otu_table(as.matrix(amr.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(amr.tbl[60:62])), sample_data(metadata.log))

# Returns an phyloseq-object containing relative abundances instead of raw read counts (uses total sum scaling).
amr.psq <- to_RA(amr.psq)

# Create microeco object
amr.mco <- file2meco::phyloseq2meco(amr.psq, auto_tidy = FALSE)
amr.mco$tidy_dataset()

### STEP 1d: Add abundance and beta diversity estimates
amr.mco$cal_abund() # Calculate the taxa abundance at each taxonomic rank using cal_abund()
amr.mco$cal_alphadiv(PD = FALSE) # Calculate alpha diversity
amr.mco$cal_betadiv(unifrac = FALSE) # Calculate beta diversity

# Save files as RDS
saveRDS(amr.psq, "out/amr.psq.RDS")
saveRDS(amr.mco, "out/amr.mco.RDS")

# Import counts annotation data from Excel file and define row names
crd.tbl <- read_excel("data/data.xlsx", sheet = "card"); crd.tbl <- column_to_rownames(crd.tbl, loc = 1)
r50.tbl <- read_excel("data/data.xlsx", sheet = "resfinder-50"); r50.tbl <- column_to_rownames(r50.tbl, loc = 1)
r80.tbl <- read_excel("data/data.xlsx", sheet = "resfinder-80"); r80.tbl <- column_to_rownames(r80.tbl, loc = 1)
meg.tbl <- read_excel("data/data.xlsx", sheet = "megares"); meg.tbl <- column_to_rownames(meg.tbl, loc = 1)
rfg.tbl <- read_excel("data/data.xlsx", sheet = "resfinderfg"); rfg.tbl <- column_to_rownames(rfg.tbl, loc = 1)
sar.tbl <- read_excel("data/data.xlsx", sheet = "sarg"); sar.tbl <- column_to_rownames(sar.tbl, loc = 1)
pan.tbl <- read_excel("data/data.xlsx", sheet = "panres"); pan.tbl <- column_to_rownames(pan.tbl, loc = 1)

# Remove decimals from columns in data frames
crd.tbl[1:59] <- lapply(crd.tbl, as.integer)
r50.tbl[1:59] <- lapply(r50.tbl, as.integer)
r80.tbl[1:59] <- lapply(r80.tbl, as.integer)
meg.tbl[1:59] <- lapply(meg.tbl, as.integer)
rfg.tbl[1:59] <- lapply(rfg.tbl, as.integer)
sar.tbl[1:59] <- lapply(sar.tbl, as.integer)
pan.tbl[1:59] <- lapply(pan.tbl, as.integer)

### STEP 1b: Create phyloseq object
crd.psq <- phyloseq(otu_table(as.matrix(crd.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(crd.tbl[60:63])), sample_data(metadata.log))
r50.psq <- phyloseq(otu_table(as.matrix(r50.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(r50.tbl[60])), sample_data(metadata.log))
r80.psq <- phyloseq(otu_table(as.matrix(r80.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(r80.tbl[60])), sample_data(metadata.log))
meg.psq <- phyloseq(otu_table(as.matrix(meg.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(meg.tbl[60:61])), sample_data(metadata.log))
rfg.psq <- phyloseq(otu_table(as.matrix(rfg.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(rfg.tbl[60])), sample_data(metadata.log))
sar.psq <- phyloseq(otu_table(as.matrix(sar.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(sar.tbl[60])), sample_data(metadata.log))
pan.psq <- phyloseq(otu_table(as.matrix(pan.tbl[1:59]), taxa_are_rows = TRUE), tax_table(as.matrix(pan.tbl[60])), sample_data(metadata.log))

### STEP 1c: Create microtable object for microeco downstream analyses

# Create microeco object
crd.mco <- file2meco::phyloseq2meco(crd.psq, auto_tidy = FALSE); crd.mco$tidy_dataset()
r50.mco <- file2meco::phyloseq2meco(r50.psq, auto_tidy = FALSE); r50.mco$tidy_dataset()
r80.mco <- file2meco::phyloseq2meco(r80.psq, auto_tidy = FALSE); r80.mco$tidy_dataset()
meg.mco <- file2meco::phyloseq2meco(meg.psq, auto_tidy = FALSE); meg.mco$tidy_dataset()
rfg.mco <- file2meco::phyloseq2meco(rfg.psq, auto_tidy = FALSE); rfg.mco$tidy_dataset()
sar.mco <- file2meco::phyloseq2meco(sar.psq, auto_tidy = FALSE); sar.mco$tidy_dataset()
pan.mco <- file2meco::phyloseq2meco(pan.psq, auto_tidy = FALSE); pan.mco$tidy_dataset()

### STEP 1d: Add abundance and beta diversity estimates

# Calculate the taxa abundance at each taxonomic rank using cal_abund()
crd.mco$cal_abund()
r50.mco$cal_abund()
r80.mco$cal_abund()
meg.mco$cal_abund()
rfg.mco$cal_abund()
sar.mco$cal_abund()
pan.mco$cal_abund()

# Calculate alpha diversity
crd.mco$cal_alphadiv(PD = FALSE)
r50.mco$cal_alphadiv(PD = FALSE)
r80.mco$cal_alphadiv(PD = FALSE)
meg.mco$cal_alphadiv(PD = FALSE)
rfg.mco$cal_alphadiv(PD = FALSE)
sar.mco$cal_alphadiv(PD = FALSE)
pan.mco$cal_alphadiv(PD = FALSE)

# Calculate beta diversity
crd.mco$cal_betadiv(unifrac = FALSE)
r50.mco$cal_betadiv(unifrac = FALSE)
r80.mco$cal_betadiv(unifrac = FALSE)
meg.mco$cal_betadiv(unifrac = FALSE)
rfg.mco$cal_betadiv(unifrac = FALSE)
sar.mco$cal_betadiv(unifrac = FALSE)
pan.mco$cal_betadiv(unifrac = FALSE)

# Save files as RDS
saveRDS(crd.psq, "out/crd.psq.RDS")
saveRDS(r50.psq, "out/r50.psq.RDS")
saveRDS(r80.psq, "out/r80.psq.RDS")
saveRDS(meg.psq, "out/meg.psq.RDS")
saveRDS(rfg.psq, "out/rfg.psq.RDS")
saveRDS(sar.psq, "out/sar.psq.RDS")
saveRDS(pan.psq, "out/pan.psq.RDS")

saveRDS(crd.mco, "out/crd.mco.RDS")
saveRDS(r50.mco, "out/r50.mco.RDS")
saveRDS(r80.mco, "out/r80.mco.RDS")
saveRDS(meg.mco, "out/meg.mco.RDS")
saveRDS(rfg.mco, "out/rfg.mco.RDS")
saveRDS(sar.mco, "out/sar.mco.RDS")
saveRDS(pan.mco, "out/pan.mco.RDS")
```

### Step XX: Visulize composition

```{r}
dataset <- clone(pan.mco)
dataset$sample_table$Region %<>% factor(., levels = rev(c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic")))

t1 <- trans_abund$new(dataset = dataset, taxrank = "Type", ntaxa = 8, groupmean = "Region")
C0 <- t1$plot_bar(clustering_plot = TRUE, color_values = rev(brewer.pal(n = 8, name = "Reds")))
C0[[1]] <- C0[[1]] + theme_bw() + theme(legend.position = "none", axis.title.x = element_text(size = 10), axis.ticks.y = element_blank(), axis.line.y = element_blank()) +
      ylab("ARG Relative abundance (%)") + ggtitle("PanRes*")
C0
```

```{r}
# Load the ggplot2 library
library(ggplot2)

# Plot number of identifications
data <- data.frame(
  Method = c("DeepARG", "AMR++", "CARD", "ResFinder (80)", "ResFinder (50)", "ResFinderFG", "MegaRes", "SARG", "PanRes"),
  Annotations = c(2042, 686, 56, 11, 526, 7, 321, 137, 1656))

# Create the bar plot
#ggplot(data, aes(x = factor(Method, levels = c("DeepARG", "AMR++", "CARD", "ResFinder (80)", "ResFinder (50)", "ResFinderFG", "MegaRes", "SARG", "PanRes")), y = Annotations)) +
D1 <- ggplot(data, aes(x = reorder(Method, -Annotations), y = Annotations)) +
      geom_bar(stat = "identity") + geom_text(aes(label = Annotations), vjust = 0.5) + theme_classic() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank(), axis.title.y = element_blank()) +
      labs(title = "No. of Annotations")

# Plot number of identifications
data <- data.frame(
  Method = c("***DeepARG", "**AMR++", "***CARD",
             "ResFinder (80)", "***ResFinder (50)", "***ResFinderFG",
             "***MegaRes", "***SARG", "***PanRes"),
  Annotations = c(0.32, 0.29, 0.24, 0.15, 0.18, 0.31, 0.15, 0.19, 0.20))

# Create the bar plot
D2 <- ggplot(data, aes(x = reorder(Method, -Annotations), y = Annotations)) +
      geom_bar(stat = "identity") + geom_text(aes(label = Annotations), vjust = 0.5) + theme_classic() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank(), axis.title.y = element_blank()) +
      labs(title = "adonis R2 (by Region)")

# Plot number of identifications
data <- data.frame(
  Method = c("DeepARG", "AMR++", "CARD",
             "ResFinder (80)", "ResFinder (50)", #"ResFinderFG",
             "MegaRes", "SARG", "PanRes"),
  Annotations = c(0.24, 0.29, 0.11, 0.03, 0.07, 0.07, 0.11, 0.10))

# Create the bar plot
D3 <- ggplot(data, aes(x = reorder(Method, -Annotations), y = Annotations)) +
      geom_bar(stat = "identity") + geom_text(aes(label = Annotations), vjust = 0.5) + theme_classic() +
      theme(axis.text.x = element_text(angle = 45, hjust = 1), axis.title.x = element_blank(), axis.title.y = element_blank()) +
      labs(title = "db-RDA adj. R2")

D1 + D2 + D3
```

### Step XX: Correlation analysis of Beta diversity

```{r}
dataset <- clone(pan.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))
#dataset <- dataset$merge_taxa(taxa = "Subtype"); dataset$tidy_dataset(); dataset ## Aggregate dataset to Species-level assignment
#dataset$cal_betadiv(unifrac = FALSE); class(dataset$beta_diversity) ## Create trans_beta object and measure parameter - dataset$beta_diversity

# measure parameter must be one of names(dataset$beta_diversity)
t1 <- trans_beta$new(dataset = dataset, group = "Region", measure = "bray")
# PCoA, PCA and NMDS are available
t1$cal_ordination(method = "PCoA"); class(t1$res_ordination) # t1$res_ordination is the ordination result list
# plot the PCoA result with confidence ellipse
B1 <- t1$plot_ordination(plot_color = "Region", 
                         plot_shape = "Region", plot_type = c("chull"), ellipse_chull_alpha = 0,
                         color_values = RColorBrewer::brewer.pal(8, "Dark2"), add_sample_label = "Sample") +
                         theme_bw() + theme(panel.grid = element_blank(), legend.position = "none") +
                         ggalt::geom_encircle(aes(s_shape = 1, expand = 0,
                                                  group = Region, fill = Region, colour = Region), alpha = 0.5) +
                         ggtitle("PanRes")
                         #annotate("text", x = -0.15, y = -0.3, label = "Bothnian Bay", colour = "#1B9E77") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Bothnian Sea", colour = "#D95F02") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Stockholm", colour = "#7570B3") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Sörmland", colour = "#E7298A") +
                         #annotate("text", x = -0.15, y = -0.08, label = "Dead Zone (Sörmland)", colour = "#66A61E") +
                         #annotate("text", x = -0.2, y = -0.2, label = "Östergötland", colour = "#E6AB02") +
                         #annotate("text", x = -0.2, y = 0, label = "Dead Zone (Mid-South)", colour = "#A6761D") +
                         #annotate("text", x = -0.3, y = 0.1, label = "Southern Baltic", colour = "#666666")
#B1

# manova for specified group set: such as "Region"
t1$cal_manova(); t1$res_manova

t1$cal_anosim(group = "Region"); t1$res_anosim
t1$cal_betadisper(); t1$res_betadisper

# calculate and plot sample distances within groups
t1$cal_group_distance(within_group = TRUE)
# return t1$res_group_distance
# perform ANOVA
t1$cal_group_distance_diff(method = "anova"); t1$res_group_distance_diff
# plot_group_order parameter can be used to adjust orders in x axis
P1 <- t1$plot_group_distance(boxplot_add = "mean", plot_group_order = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic")) +
      geom_boxplot(fill = brewer.pal(n = 8, name = "Dark2"), color = "black", alpha = 1, outlier.shape = NA) + theme_bw() +
      #geom_smooth(method = loess, se = TRUE, aes(group = 1)) +
      theme(legend.position = "top", axis.text.y = element_text(size = 10, angle = 0), 
            panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
            axis.text.x = element_text(size = 10, angle = 40, hjust = 1)) +  ggtitle("PanRes")

P1
```

```{r}
library("vegan")

# Extract beta diversity matrices (Bray-Curtis) for each dataset
arg.bc <- arg.mco$beta_diversity$bray
amr.bc <- amr.mco$beta_diversity$bray
crd.bc <- crd.mco$beta_diversity$bray
r80.bc <- r80.mco$beta_diversity$bray
r50.bc <- r50.mco$beta_diversity$bray
meg.bc <- meg.mco$beta_diversity$bray
rfg.bc <- rfg.mco$beta_diversity$bray
sar.bc <- sar.mco$beta_diversity$bray
pan.bc <- pan.mco$beta_diversity$bray

# Hellinger transformation for PCA
arg.bc.hel <- decostand(arg.bc, method = "hellinger"); arg.bc.hel.pca <- rda(arg.bc.hel)
amr.bc.hel <- decostand(amr.bc, method = "hellinger"); amr.bc.hel.pca <- rda(amr.bc.hel)
crd.bc.hel <- decostand(crd.bc, method = "hellinger"); crd.bc.hel.pca <- rda(crd.bc.hel)
r80.bc.hel <- decostand(r80.bc, method = "hellinger"); r80.bc.hel.pca <- rda(r80.bc.hel)
r50.bc.hel <- decostand(r50.bc, method = "hellinger"); r50.bc.hel.pca <- rda(r50.bc.hel)
meg.bc.hel <- decostand(meg.bc, method = "hellinger"); meg.bc.hel.pca <- rda(meg.bc.hel)
rfg.bc.hel <- decostand(rfg.bc, method = "hellinger"); rfg.bc.hel.pca <- rda(rfg.bc.hel)
sar.bc.hel <- decostand(sar.bc, method = "hellinger"); sar.bc.hel.pca <- rda(sar.bc.hel)
pan.bc.hel <- decostand(pan.bc, method = "hellinger"); pan.bc.hel.pca <- rda(pan.bc.hel)

# Test between ARG and OTU
arg.amr.pro <- procrustes(X = arg.bc.hel.pca, Y = amr.bc.hel.pca, symmetric = TRUE); summary(arg.amr.pro)
arg.amr.prt <- protest(X = arg.bc.hel.pca, Y = amr.bc.hel.pca, scores = "sites", permutations = 999)
arg.amr.prt
arg.crd.pro <- procrustes(X = arg.bc.hel.pca, Y = crd.bc.hel.pca, symmetric = TRUE); summary(arg.crd.pro)
arg.crd.prt <- protest(X = arg.bc.hel.pca, Y = crd.bc.hel.pca, scores = "sites", permutations = 999)
arg.crd.prt
arg.r80.pro <- procrustes(X = arg.bc.hel.pca, Y = r80.bc.hel.pca, symmetric = TRUE); summary(arg.r80.pro)
arg.r80.prt <- protest(X = arg.bc.hel.pca, Y = r80.bc.hel.pca, scores = "sites", permutations = 999)
arg.r80.prt
arg.r50.pro <- procrustes(X = arg.bc.hel.pca, Y = r50.bc.hel.pca, symmetric = TRUE); summary(arg.r50.pro)
arg.r50.prt <- protest(X = arg.bc.hel.pca, Y = r50.bc.hel.pca, scores = "sites", permutations = 999)
arg.r50.prt
arg.meg.pro <- procrustes(X = arg.bc.hel.pca, Y = meg.bc.hel.pca, symmetric = TRUE); summary(arg.meg.pro)
arg.meg.prt <- protest(X = arg.bc.hel.pca, Y = meg.bc.hel.pca, scores = "sites", permutations = 999)
arg.meg.prt
arg.rfg.pro <- procrustes(X = arg.bc.hel.pca, Y = rfg.bc.hel.pca, symmetric = TRUE); summary(arg.rfg.pro)
arg.rfg.prt <- protest(X = arg.bc.hel.pca, Y = rfg.bc.hel.pca, scores = "sites", permutations = 999)
arg.rfg.prt
arg.sar.pro <- procrustes(X = arg.bc.hel.pca, Y = sar.bc.hel.pca, symmetric = TRUE); summary(arg.sar.pro)
arg.sar.prt <- protest(X = arg.bc.hel.pca, Y = sar.bc.hel.pca, scores = "sites", permutations = 999)
arg.sar.prt
arg.pan.pro <- procrustes(X = arg.bc.hel.pca, Y = pan.bc.hel.pca, symmetric = TRUE); summary(arg.pan.pro)
arg.pan.prt <- protest(X = arg.bc.hel.pca, Y = pan.bc.hel.pca, scores = "sites", permutations = 999)
arg.pan.prt

#plot(arg.amr.pro, kind = 1)
plot(arg.amr.pro, kind = 1, main = "DeepARG & AMR++", ar.col = "gray")
lines(arg.amr.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.amr.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.amr.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")

#plot(arg.crd.pro, kind = 1)
plot(arg.crd.pro, kind = 1, main = "DeepARG & CARD", ar.col = "gray")
lines(arg.crd.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.crd.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.crd.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")

#plot(arg.r80.pro, kind = 1)
plot(arg.r80.pro, kind = 1, main = "DeepARG & ResFinder (80%)", ar.col = "gray")
lines(arg.r80.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.r80.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.r80.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")

#plot(arg.r50.pro, kind = 1)
plot(arg.r50.pro, kind = 1, main = "DeepARG & ResFinder (50%)", ar.col = "gray")
lines(arg.r50.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.r50.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.r50.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")

#plot(arg.meg.pro, kind = 1)
plot(arg.meg.pro, kind = 1, main = "DeepARG & MegaRes", ar.col = "gray")
lines(arg.meg.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.meg.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.meg.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")

#plot(arg.rfg.pro, kind = 1)
plot(arg.rfg.pro, kind = 1, main = "DeepARG & ResFinderFG", ar.col = "gray")
lines(arg.rfg.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.rfg.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.rfg.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")

#plot(arg.sar.pro, kind = 1)
plot(arg.sar.pro, kind = 1, main = "DeepARG & SARG", ar.col = "gray")
lines(arg.sar.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.sar.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.sar.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")

#plot(arg.pan.pro, kind = 1)
plot(arg.pan.pro, kind = 1, main = "DeepARG & PanRes", ar.col = "gray")
lines(arg.pan.pro, type = "segments", col = "gray", lwd = 2, lty = 3)
points(arg.pan.pro, display = c("target"), col = "black", cex = 1, pch = 21, bg = "#8DA0CB")
points(arg.pan.pro, display = c("rotated"), col = "black", cex = 1, pch = 21, bg = "#FDAE61")
```

### Step XX: Correlation analysis of Alpha diversity

```{r}
# Pearson Correlation of Chao1 Richness

# Combine your matrices into a single data frame
df <- cbind(ARG = (arg.mco$alpha_diversity$Pielou),
            AMR = (amr.mco$alpha_diversity$Pielou),
            CRD = (crd.mco$alpha_diversity$Pielou),
            #R80 = (r80.mco$alpha_diversity$Pielou),
            R50 = (r50.mco$alpha_diversity$Pielou),
            MEG = (meg.mco$alpha_diversity$Pielou),
            #RFG = (rfg.mco$alpha_diversity$Pielou),
            SAR = (sar.mco$alpha_diversity$Pielou),
            PAN = (pan.mco$alpha_diversity$Pielou))

library(reshape2)

#calculate correlation coefficients, rounded to 2 decimal places
cor_df <- round(cor(df), 2)

#melt the data frame
melted_cor <- melt(cor_df)

#view head of melted data frame
head(melted_cor)

library(ggplot2)

#create correlation heatmap
ggplot(data = melted_cor, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() + ggtitle("Pielou Evenness") +
  geom_text(aes(Var2, Var1, label = value), size = 3) +
  scale_fill_gradient2(low = "blue", high = "red", limit = c(-1,1), name="Pearson\nCorrelation") +
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), panel.background = element_blank())
```

### Step XX: Environmental Association

```{r}
# Clone the dataset and set Region as a factor with specified levels
dataset <- clone(arg.mco)
dataset$sample_table$Region <- factor(dataset$sample_table$Region, levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", 
                                                                             "Dead Zone (Sörmland)", "Östergötland", 
                                                                             "Dead Zone (Mid-South)", "Southern Baltic"))

# Rename columns with special characters
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Depth")] <- "Depth*"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Salinity")] <- "Salinity***"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Temperature")] <- "Temperature**"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "O2")] <- "Dissolved O2***"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "C.N.Ratio")] <- "C:N Ratio**"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Delta.15N")] <- "δ15N***"
colnames(dataset$sample_table)[which(names(dataset$sample_table) == "Delta.13C")] <- "δ13C*"

# Initialize trans_env object and add environmental data
t1 <- trans_env$new(dataset = dataset, add_data = dataset$sample_table[, c(7:10, 13:15)])

# Perform dbRDA analysis using Bray-Curtis distance
set.seed(0327)
t1$cal_ordination(method = "dbRDA", use_measure = "bray")

# Transform ordination results for visualization
t1$trans_ordination(adjust_arrow_length = TRUE, max_perc_env = 1.5)

# Plot ordination with environmental variables and color by Region
t1$plot_ordination(plot_color = "Region", plot_type = "point", color_values = RColorBrewer::brewer.pal(8, "Dark2")) +
                   theme_bw() + theme(panel.grid = element_blank(), legend.position = "none")  +  ggtitle("PanRes")

# View ordination terms
t1$res_ordination_terms

# Perform ANOVA on ordination axes
t1$cal_ordination_anova()
t1$res_ordination_axis

# Perform envfit analysis
t1$cal_ordination_envfit()
t1$res_ordination_envfit

t1$res_ordination_R2
```

```{r}
dataset <- clone(arg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D1 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#8DA0CB", lm_equation = FALSE, line_color = "#8DA0CB",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#8DA0CB", alpha = 0.05) +  
                         ggtitle("DeepARG")

dataset <- clone(amr.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D2 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("AMR++")

dataset <- clone(crd.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D3 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("CARD")

dataset <- clone(r80.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D4 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("ResFinder (80)")

dataset <- clone(r50.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D5 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("ResFinder (50)")

dataset <- clone(rfg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D6 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("ResFinder FG")

dataset <- clone(meg.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D7 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("MegaRes")

dataset <- clone(sar.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D8 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("SARG")

dataset <- clone(pan.mco)
dataset$sample_table$Region %<>% factor(., levels = c("Bothnian Bay", "Bothnian Sea", "Stockholm", "Sörmland", "Dead Zone (Sörmland)", "Östergötland", "Dead Zone (Mid-South)","Southern Baltic"))

## Assess the relationship between environmental factors and alpha diversity, please use add_abund_table parameter in the cal_cor function. The correlation result is stored in object$res_cor ...
t1 <- trans_env$new(dataset = dataset, env_cols = c(1:4), complete_na = TRUE) ## env_cols is used to add the environmental data
D9 <- t1$plot_scatterfit(x = "Distance", y = dataset$beta_diversity$bray[rownames(t1$data_env), rownames(t1$data_env)], 
                         type = "lm", point_size = 2, point_alpha = 0, label.x.npc = "left", label.y.npc = "top", 
                         line_alpha = 0.25, line_se_color = "#FDAE61", lm_equation = FALSE, line_color = "#FDAE61",
                         x_axis_title = "km distance", y_axis_title = "Bray-Curtis distance") + theme_bw() + 
                         theme(panel.grid = element_blank()) + geom_point(colour = "#FDAE61", alpha = 0.05) +  
                         ggtitle("PanRes")

D1 + D2 + D3 + D4 + D5 + D6 + D7 + D8 + D9
```

